---
title: 'Homework 4: Machine Learning'
author: "Jamie McGraw"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  html_document:
    theme: flatly
    highlight: zenburn
    number_sections: yes
    toc: yes
    toc_float: yes
    code_folding: show
---

```{r}
#| label: load-libraries
#| echo: false # This option disables the printing of code (only output is displayed).
#| message: false
#| warning: false
options(scipen = 999) #disable scientific notation
library(tidyverse)
library(tidymodels)
library(GGally)
library(sf)
library(leaflet)
library(janitor)
library(rpart.plot)
library(here)
library(scales)
library(vip)
library(C50)
library(GGally)
```


# The Bechdel Test

https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/

The [Bechdel test](https://bechdeltest.com) is a way to assess how women are depicted in Hollywood movies.  In order for a movie to pass the test:

1. It has to have at least two [named] women in it
2. Who talk to each other
3. About something besides a man

There is a nice article and analysis you can find here https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/
We have a sample of 1394 movies and we want to fit a model to predict whether a film passes the test or not.

```{r read_data}

bechdel <- read_csv(here::here("data", "bechdel.csv")) %>% 
  mutate(test = factor(test)) 
glimpse(bechdel)

```
How many films fail/pass the test, both as a number and as a %?

```{r}

#55% of movies fail the test (772)
bechdel %>% 
  group_by(test) %>% 
  summarize(count = n()) %>% 
  mutate(percent = count/sum(count)*100)

```


## Movie scores
```{r}
ggplot(data = bechdel, aes(
  x = metascore,
  y = imdb_rating,
  colour = test
)) +
  geom_point(alpha = .3, size = 3) +
  scale_colour_manual(values = c("tomato", "olivedrab")) +
  labs(
    x = "Metacritic score",
    y = "IMDB rating",
    colour = "Bechdel test"
  ) +
 theme_light()
```


# Split the data
```{r}
# **Split the data**

set.seed(123)

data_split <- initial_split(bechdel, # updated data
                           prop = 0.8, 
                           strata = test)

bechdel_train <- training(data_split) 
bechdel_test <- testing(data_split)
```

Check the counts and % (proportions) of the `test` variable in each set.
```{r}
#repeat above code for different datasets
bechdel_train %>% 
  group_by(test) %>% 
  summarise(count = n()) %>% 
  mutate(percent = count/sum(count))


bechdel_test %>% 
  group_by(test) %>% 
  summarise(count = n()) %>% 
  mutate(percent = count/sum(count))

#so it looks like the samples are well balanced. (both within a percent of each other)
```

## Feature exploration

## Any outliers? 

```{r}

bechdel %>% 
  select(test, budget_2013, domgross_2013, intgross_2013, imdb_rating, metascore) %>% 

    pivot_longer(cols = 2:6,
               names_to = "feature",
               values_to = "value") %>% 
  ggplot()+
  aes(x=test, y = value, fill = test)+
  coord_flip()+
  geom_boxplot()+
  facet_wrap(~feature, scales = "free")+
  theme_bw()+
  theme(legend.position = "none")+
  labs(x=NULL,y = NULL)

```

## Scatterplot - Correlation Matrix

Write a paragraph discussing the output of the following 
```{r, warning=FALSE, message=FALSE}
bechdel %>% 
  select(test, budget_2013, domgross_2013, intgross_2013, imdb_rating, metascore)%>% 
  ggpairs(aes(colour=test), alpha=0.2)+
  theme_bw()

#Noteworthy Differences: Slightly more movies pass the Bechdel test and these films tend to have higher budgets. This is relevant because there's a strong correlation between a movie's budget and its gross earnings both domestically and internationally. Interestingly, this correlation is slightly stronger for movies that pass the test, implying that investing in such movies may offer better returns. 

#Less Significant Differences: The data shows a right skew in movies' performance, indicating blockbuster hits as potential outliers that might influence perceptions of overall success. However, variations in revenue, both domestically and internationally, aren't significantly large, nor are differences in correlations. Despite some noticeable correlation differences, their absolute values are low, undermining the assertion of meaningful relationships. The Bechdel test does not appear to significantly affect movie ratings.
```


## Categorical variables

Write a paragraph discussing the output of the following 
```{r}
bechdel %>% 
  group_by(genre, test) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))
#The analysis indicates that movie genre significantly influences Bechdel test results, with substantial variation among genres. Action films, the most popular genre, predominantly fail the test, possibly reflecting gender representation within this genre. The other two popular genres, Comedy and Drama, slightly lean towards passing the test. Removing action films might provide additional insights. Genres like Animation and Crime show a male-oriented skew, while Horror tends to pass. However, with the number of films being heavily concentrated in a few genres, we should consider if the proportions are truly representative, particularly for less popular genres. This may limit the conclusions we can draw from less populated genres.  
 
bechdel %>% 
  group_by(rated, test) %>%
  summarise(n = n()) %>% 
  mutate(prop = n/sum(n))

#Film ratings do seem to vary with Bechdel test outcomes, although perhaps less materially than with genres. Most strikingly, G and NC-17 rated films, albeit smaller categories, lean heavily towards failing the test, with approximately 62% and 83% failure rates respectively. More mainstream categories like PG, PG-13, and R present a more balanced distribution, although they slightly skew towards failing. The R-rating, typically the largest category, shows a notable 57% failure rate, suggesting a broader industry trend towards less female interaction in dialogues. Even G-rated children's films largely fail the Bechdel test, raising questions about gender representation from early cinematic exposure.
```

# Train first models. `test ~ metascore + imdb_rating`

```{r}
lr_mod <- logistic_reg() %>% 
  set_engine(engine = "glm") %>% 
  set_mode("classification")

lr_mod


tree_mod <- decision_tree() %>% 
  set_engine(engine = "C5.0") %>% 
  set_mode("classification")

tree_mod 
```

```{r}


lr_fit <- lr_mod %>% # parsnip model
  fit(test ~ metascore + imdb_rating, # a formula
    data = bechdel_train # dataframe
  )

tree_fit <- tree_mod %>% # parsnip model
  fit(test ~ metascore + imdb_rating, # a formula
    data = bechdel_train # dataframe
  )
```

## Logistic regression

```{r}
lr_fit %>%
  broom::tidy()

lr_preds <- lr_fit %>%
  augment(new_data = bechdel_train) %>%
  mutate(.pred_match = if_else(test == .pred_class, 1, 0))

```

### Confusion matrix

```{r}
lr_preds %>% 
  conf_mat(truth = test, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")


```


## Decision Tree
```{r}
tree_preds <- tree_fit %>%
  augment(new_data = bechdel) %>%
  mutate(.pred_match = if_else(test == .pred_class, 1, 0)) 


```

```{r}
tree_preds %>% 
  conf_mat(truth = test, estimate = .pred_class) %>% 
  autoplot(type = "heatmap")
```

## Draw the decision tree

```{r}
draw_tree <- 
    rpart::rpart(
        test ~ metascore + imdb_rating,
        data = bechdel_train, # uses data that contains both birth weight and `low`
        control = rpart::rpart.control(maxdepth = 5, cp = 0, minsplit = 10)
    ) %>% 
    partykit::as.party()
plot(draw_tree)

```

# Cross Validation

Run the code below. What does it return?

```{r}
set.seed(123)
bechdel_folds <- vfold_cv(data = bechdel_train, 
                          v = 10, 
                          strata = test)
bechdel_folds

#tibble with folds
```

## `fit_resamples()`

Trains and tests a resampled model.

```{r}
lr_fit <- lr_mod %>%
  fit_resamples(
    test ~ metascore + imdb_rating,
    resamples = bechdel_folds
  )


tree_fit <- tree_mod %>%
  fit_resamples(
    test ~ metascore + imdb_rating,
    resamples = bechdel_folds
  )
```


## `collect_metrics()`

Unnest the metrics column from a tidymodels `fit_resamples()`
```{r}

collect_metrics(lr_fit)
collect_metrics(tree_fit)


```


```{r}
tree_preds <- tree_mod %>% 
  fit_resamples(
    test ~ metascore + imdb_rating, 
    resamples = bechdel_folds,
    control = control_resamples(save_pred = TRUE) #<<
  )

# What does the data for ROC look like?
tree_preds %>% 
  collect_predictions() %>% 
  roc_curve(truth = test, .pred_Fail)  

# Draw the ROC
tree_preds %>% 
  collect_predictions() %>% 
  roc_curve(truth = test, .pred_Fail) %>% 
  autoplot()

#The curve only appears to perform slighlty better than chance, and at the top-right even slightly underperforms --> not very good

```


# Build a better training set with `recipes`

## Preprocessing options

- Encode categorical predictors
- Center and scale variables
- Handle class imbalance
- Impute missing data
- Perform dimensionality reduction 
- ... ...

## To build a recipe

1. Start the `recipe()`
1. Define the variables involved
1. Describe **prep**rocessing [step-by-step]

## Collapse Some Categorical Levels

Do we have any `genre` with few observations?  Assign genres that have less than 3% to a new category 'Other'


```{r}
#| echo = FALSE
bechdel %>% 
  count(genre) %>% 
  mutate(genre = fct_reorder(genre, n)) %>% 
  ggplot(aes(x = genre, 
             y = n)) +
  geom_col(alpha = .8) +
  coord_flip() +
  labs(x = NULL) +
  geom_hline(yintercept = (nrow(bechdel_train)*.03), lty = 3)+
  theme_light()
```


```{r}
movie_rec <-
  recipe(test ~ .,
         data = bechdel_train) %>%
  
  # Genres with less than 5% will be in a category 'Other'
    step_other(genre, threshold = .03) 
```
  

## Before recipe

```{r}
#| echo = FALSE
bechdel_train %>% 
  count(genre, sort = TRUE)
```


## After recipe

```{r}
movie_rec %>% 
  prep() %>% 
  bake(new_data = bechdel_train) %>% 
  count(genre, sort = TRUE)

#So biography is now largest standalone genre
```

## `step_dummy()`

Converts nominal data into numeric dummy variables

```{r}
#| results = "hide"
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_dummy(all_nominal_predictors()) 

movie_rec 
```

## Let's think about the modelling 

What if there were no films with `rated` NC-17 in the training data?

-  Will the model have a coefficient for `rated` NC-17?
-  What will happen if the test data includes a film with `rated` NC-17?

## `step_novel()`

Adds a catch-all level to a factor for any new values not encountered in model training, which lets R intelligently predict new levels in the test set.

```{r}

movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal_predictors) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal_predictors()) 

```


## `step_zv()`

Intelligently handles zero variance variables (variables that contain only a single value)

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes()) 
  
```


## `step_normalize()`

Centers then scales numeric variable (mean = 0, sd = 1)

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes())  %>% 
  step_normalize(all_numeric()) 

```


## `step_corr()`

Removes highly correlated variables

```{r}
movie_rec <- recipe(test ~ ., data = bechdel) %>%
  step_other(genre, threshold = .03) %>% 
  step_novel(all_nominal(), -all_outcomes()) %>% # Use *before* `step_dummy()` so new level is dummified
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_zv(all_numeric(), -all_outcomes())  %>% 
  step_normalize(all_numeric()) %>% 
  step_corr(all_predictors(), threshold = 0.75, method = "spearman") 



movie_rec
```


# Define different models to fit

```{r}
## Model Building

# 1. Pick a `model type`
# 2. set the `engine`
# 3. Set the `mode`: regression or classification

# Logistic regression
log_spec <-  logistic_reg() %>%  # model type
  set_engine(engine = "glm") %>%  # model engine
  set_mode("classification") # model mode

# Show your model specification
log_spec

# Decision Tree
tree_spec <- decision_tree() %>%
  set_engine(engine = "C5.0") %>%
  set_mode("classification")

tree_spec

# Random Forest
library(ranger)

rf_spec <- 
  rand_forest() %>% 
  set_engine("ranger", importance = "impurity") %>% 
  set_mode("classification")


# Boosted tree (XGBoost)
library(xgboost)

xgb_spec <- 
  boost_tree() %>% 
  set_engine("xgboost") %>% 
  set_mode("classification") 

# K-nearest neighbour (k-NN)
knn_spec <- 
  nearest_neighbor(neighbors = 4) %>% # we can adjust the number of neighbors 
  set_engine("kknn") %>% 
  set_mode("classification") 
```


# Bundle recipe and model with `workflows`


```{r}
log_wflow <- # new workflow object
 workflow() %>% # use workflow function
 add_recipe(movie_rec) %>%   # use the new recipe
 add_model(log_spec)   # add your model spec

# show object
log_wflow


## A few more workflows

tree_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(tree_spec) 

rf_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(rf_spec) 

xgb_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(xgb_spec)

knn_wflow <-
 workflow() %>%
 add_recipe(movie_rec) %>% 
 add_model(knn_spec)

```

HEADS UP

1. How many models have you specified?
2. What's the difference between a model specification and a workflow?
3. Do you need to add a formula (e.g., `test ~ .`)  if you have a recipe?
```{r}
#1. we've specified 5 models
#2. a workflow is a function that can take recipe/prepocessing specifications and model specifications, 
#... this is what allowed us to use the same movie_rec recipe with all the different model specifications (knn, logisitc, r-forest etc..) so we can evaluate these models on the same prepocessed data
```


# Model Comparison

You now have all your models. Adapt the code from slides `code-from-slides-CA-housing.R`, line 400 onwards to assess which model gives you the best classification. 


```{r}
## Evaluate Models
#use bechdel folds and adjust conf_mat variable = test

###################MODEL 1: LOGISTIC REGRESSION########################### --> this model does particularly poorly
###################MODEL 1: LOGISTIC REGRESSION###########################
log_res <- log_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, accuracy,
      kap, roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)) 

# Show average performance over all folds (note that we use log_res):
log_res %>%  collect_metrics(summarize = TRUE)

# Show performance for every single fold:
log_res %>%  collect_metrics(summarize = FALSE)


## `collect_predictions()` and get confusion matrix{.smaller}

log_pred <- log_res %>% collect_predictions()

log_pred %>%  conf_mat(truth = test, estimate = .pred_class) #adjust confusion matrix code

log_pred %>% 
  conf_mat(test, .pred_class) %>% 
  autoplot(type = "mosaic") +
  geom_label(aes(
      x = (xmax + xmin) / 2, 
      y = (ymax + ymin) / 2, 
      label = c("TP", "FN", "FP", "TN")))

log_pred %>% 
  conf_mat(test, .pred_class) %>% 
  autoplot(type = "heatmap")


## ROC Curve

log_pred %>% 
  group_by(id) %>% # id contains our folds
  roc_curve(truth = test, .pred_Fail) %>% 
  autoplot()

###################MODEL 2: DECISION TREE###########################
###################MODEL 2: DECISION TREE###########################

tree_res <-
  tree_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

tree_res %>%  collect_metrics(summarize = TRUE)

###################MODEL 3: RANDOM FOREST###########################
###################MODEL 3: RANDOM FOREST###########################


rf_res <-
  rf_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

rf_res %>%  collect_metrics(summarize = TRUE)

###################MODEL 4: XGBOOST###########################
###################MODEL 4: XGBOOST###########################

xgb_res <- 
  xgb_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

xgb_res %>% collect_metrics(summarize = TRUE)

###################MODEL 5: K-NEAREST NEIGHBOUR###########################
###################MODEL 5: K-NEAREST NEIGHBOUR###########################

knn_res <- 
  knn_wflow %>% 
  fit_resamples(
    resamples = bechdel_folds, 
    metrics = metric_set(
      recall, precision, f_meas, 
      accuracy, kap,
      roc_auc, sens, spec),
    control = control_resamples(save_pred = TRUE)
    ) 

knn_res %>% collect_metrics(summarize = TRUE)


###################MODEL COMPARISON SETUP (ROC MEAN CALC)###########################
###################MODEL COMPARISON SETUP (ROC MEAN CALC)###########################

log_metrics <- 
  log_res %>% 
  collect_metrics(summarise = TRUE) %>%
  # add the name of the model to every row
  mutate(model = "Logistic Regression") 

tree_metrics <- 
  tree_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Decision Tree")

rf_metrics <- 
  rf_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Random Forest")

xgb_metrics <- 
  xgb_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "XGBoost")

knn_metrics <- 
  knn_res %>% 
  collect_metrics(summarise = TRUE) %>%
  mutate(model = "Knn")

# create dataframe with all models
model_compare <- bind_rows(log_metrics,
                           tree_metrics,
                           rf_metrics,
                           xgb_metrics,
                           knn_metrics) 

#Pivot wider to create barplot
  model_comp <- model_compare %>% 
  select(model, .metric, mean, std_err) %>% 
  pivot_wider(names_from = .metric, values_from = c(mean, std_err)) 

  
###################MODEL RESULTS###########################
###################MODEL RESULTS###########################  
  
# model performance from bar chart ordered best to worst (based on mean ROC): RF, XGBoost, Decision Tree, Knn, Logistic regression
  model_comp %>% 
  arrange(mean_roc_auc) %>% 
  mutate(model = fct_reorder(model, mean_roc_auc)) %>% # order results
  ggplot(aes(model, mean_roc_auc, fill=model)) +
  geom_col() +
  coord_flip() +
  scale_fill_brewer(palette = "Blues") +
   geom_text(
     size = 3,
     aes(label = round(mean_roc_auc, 2), 
         y = mean_roc_auc + 0.08),
     vjust = 1
  )+
  theme_light()+
  theme(legend.position = "none")+
  labs(y = NULL)

#look at other metrics to confirm superiority of RF: yes it outperforms XG Boost along most metrics except mean specificity
model_comp

# use lastfit() on RF given it is the best performing model here
 
last_fit_rf <- last_fit(rf_wflow, 
                        split = data_split,
                        metrics = metric_set(
                          accuracy, f_meas, kap, precision,
                          recall, roc_auc, sens, spec))


#Compare this to training - as expected, training is marginally better
last_fit_rf %>% collect_metrics(summarize = TRUE)
rf_res %>% collect_metrics(summarize = TRUE)
```

# Final model analysis 
```{r}
######## Variable importance ###############
######## Variable importance ###############

#1. budget 2013 contributes the most
#2. imdb_rating contributes the second most, which might makes sense because of the low correlation to budget 2013 from our discussion above, perhaps adding the most independent information to the model
#3.surprising: Horror was the most important predictor for genres.

library(vip)

last_fit_rf %>% 
  pluck(".workflow", 1) %>%   
  pull_workflow_fit() %>% 
  vip(num_features = 10) +
  theme_light()


## Final Confusion Matrix - we can see the model is better at predicting fails than passes. To be expected given most data points were Fails

last_fit_rf %>%
  collect_predictions() %>% 
  conf_mat(test, .pred_class) %>% 
  autoplot(type = "heatmap")


## Final ROC curve
last_fit_rf %>% 
  collect_predictions() %>% 
  roc_curve(test, .pred_Fail) %>% 
  autoplot()



```



# Deliverables

There is a lot of explanatory text, comments, etc. You do not need these, so delete them and produce a stand-alone document that you could share with someone. Knit the edited and completed R Markdown (Rmd) file as a Word or HTML document (use the "Knit" button at the top of the script editor window) and upload it to Canvas. You must be commiting and pushing your changes to your own Github repo as you go along.

# Details

-   Who did you collaborate with: Just your lecture code
-   Approximately how much time did you spend on this problem set: 4 hours
-   What, if anything, gave you the most trouble: 

**Please seek out help when you need it,** and remember the [15-minute rule](https://dsb2023.netlify.app/syllabus/#the-15-minute-rule){target="_blank"}. You know enough R (and have enough examples of code from class and your readings) to be able to do this. If you get stuck, ask for help from others, post a question on Slack-- and remember that I am here to help too!

> As a true test to yourself, do you understand the code you submitted and are you able to explain it to someone else?

# Rubric

13/13: Problem set is 100% completed. Every question was attempted and answered, and most answers are correct. Code is well-documented (both self-documented and with additional comments as necessary). Used tidyverse, instead of base R. Graphs and tables are properly labelled. Analysis is clear and easy to follow, either because graphs are labeled clearly or you've written additional text to describe how you interpret the output. Multiple Github commits. Work is exceptional. I will not assign these often.

8/13: Problem set is 60--80% complete and most answers are correct. This is the expected level of performance. Solid effort. Hits all the elements. No clear mistakes. Easy to follow (both the code and the output). A few Github commits.

5/13: Problem set is less than 60% complete and/or most answers are incorrect. This indicates that you need to improve next time. I will hopefully not assign these often. Displays minimal effort. Doesn't complete all components. Code is poorly written and not documented. Uses the same type of plot for each graph, or doesn't use plots appropriate for the variables being analyzed. No Github commits.
